Random number generation and applications on GPU’s in R?

### 2021-10-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.Rnw
debugging
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@2903 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Now we use the streams created in section \ref{create_streams} to generate two vectors of 6 i.i.d. $U (0,1)$ random numbers using the function \fct{runif}. If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host. By using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers. 
<<simStuff2, eval=TRUE,tidy=TRUE>>=
as.vector(gpuRandom::runif(n=4, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))[-(1:6)]
@
The arguments of the \fct{gpuRandom::runif} are described as follows:
\begin{itemize}
  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
  \item \code{streams}: streams for random number generation.
  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
\end{itemize}
<<include=FALSE, tidy=TRUE>>=
sim_2 = runif(n=7,streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type="double")
as.vector(sim_2)
@
Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}. Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the \code{myStreamsGpu1} after it being used.
<<eval=TRUE,tidy=TRUE>>=
t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
@
Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart. Streams created on GPU does not remain in the memory when a new \proglang{R} session begins. However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations. Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
<<save streams,echo=TRUE,eval=TRUE>>=
saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
# Load the streams object as streams_saved
streams_saved <- vclMatrix(readRDS("myStreams.rds"))
@

>>>>

Now we use the streams created in section \ref{create_streams} to generate two vectors of 6 i.i.d. $U (0,1)$ random numbers using the function \fct{runif}. If \code{streams} and \code{Nglobal} are both missing, then by default \code{Nglobal} is set as (64,8), and a `vclMatrix' of \code{prod(Nglobal)} streams is supplied. The output is an S4 object (a `\code{vclVector}' or a `\code{vclMatrix}'), to view the generated random numbers, we need to convert it to an R vector or matrix, in this way, the random numbers are moved from the global memory of device to the host. By using \code{myStreamsGpu2} which is the same as \code{myStreamsGpu1},  \fct{runif} returns the same sequence of random numbers. 
<<simStuff2, eval=TRUE,tidy=TRUE>>=
as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu1, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
as.vector(gpuRandom::runif(n=10, streams=myStreamsGpu2, Nglobal=c(1,nrow(myStreamsGpu1)), type = "double"))
@
The arguments of the \fct{gpuRandom::runif} are described as follows:
\begin{itemize}
  \item \code{n}: a vector of length 2 specifying the row and column number if to create a matrix, or a number specifying the length if to create a vector.
  \item \code{streams}: streams for random number generation.
  \item \code{Nglobal}: global index space. Default is set as $(64,8)$.
  \item \code{type}: ``double'' or ``float'' format of generated random numbers.
\end{itemize}
Reusing \code{myStreamsGpu1} will produce a vector different from \code{sim_1}. Because each time a random number is generated, the current state of the stream is changed (advances by one position), below shows the \code{myStreamsGpu1} after it being used.
<<eval=TRUE,tidy=TRUE>>=
t(matrix(as.matrix(myStreamsGpu1), nrow(myStreamsCpu), ncol(myStreamsCpu), dimnames = dimnames(myStreamsCpu)))
@
Unlike the objects created by \proglang{R}, by default they remain in the memory after a restart. Streams created on GPU does not remain in the memory when a new \proglang{R} session begins. However, we can save the streams on the CPU in a file using the \fct{save} function, and later recall the streams with the \fct{load} function, in this way we can reproduce the same sequence of random numbers in simulations. Below is a trivial example, in which we demonstrate saving streams to a data file called \code{streams_from_GPU.rds} on CPU and then load it back and transfer it to GPU.
<<save streams,echo=TRUE,eval=TRUE>>=
saveRDS(as.matrix(createStreamsGpu(n=4)), "myStreams.rds")
# Load the streams object as streams_saved
streams_saved <- vclMatrix(readRDS("myStreams.rds"))
@
